---
title: "hw_1_zl2860"
author: "Zongchao Liu"
date: "9/15/2019"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 0.1 

```{r}
#----------STEP 1: Create a dataframe
set.seed(886) # set.seed() to ensure the reproductivity
df_1 <- tibble(
  random_sample = rnorm(8), # create a random sample of size 8 from a standard Normal distribution
  vec_logical = random_sample > 0, # create a logical vector indicating whether elements of the sample are greater than 0
  vec_character = as.character(c("a","a","a","a","a","a","a","a")), # create a character vector of length 8
  vec_factor = factor(c("good","bad","medium","good","bad","medium","bad","medium"),levels = c("good","medium","bad"))
) #create a factor vector of length 8, with 3 different factor ???levels???
levels(pull(df_1,vec_factor)) #see the levels of the created factor vector




```


```{r}
#----------STEP 2: Take the mean of each variable in the dataframe
mean(pull(df_1,random_sample)) # return numeric values
mean(pull(df_1,vec_logical)) # return numeric values
mean(pull(df_1,vec_factor)) # return NAs
mean(pull(df_1,vec_factor)) # return NAs
```
What works and what doesn???t?


```{r, eval = FALSE}
#----------STEP 3: Convert dtype & Multiplication
#STEP 3.1:  Convert dtype
as.numeric(pull(df_1,vec_logical)) # return numeric values
as.numeric(pull(df_1,vec_character)) # return NAs
as.numeric(pull(df_1,vec_factor)) # return numeric values
```
What happens, and why? Does this help explain what happens when you try to take the mean?

```{r,eval = FALSE}
#STEP 3.2: Convert dtype & Multiplication
result1 =  as.numeric(pull(df_1,vec_logical)) * pull(df_1,random_sample) # convert the logical vector to numeric, and multiply the random sample by the result
result2 = as.factor(pull(df_1,vec_logical)) * pull(df_1,random_sample) # convert the logical vector to a factor, and multiply the random sample by the result
result3 = as.numeric(as.factor(pull(df_1,vec_logical))) * pull(df_1,random_sample) # convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```

## Promblem 2 
## Create a new dataframe & Write a short description of the vectors using inline R code
```{r}
set.seed(886)
#Create a data frame 
df_2 <- tibble(
  x = rnorm(500), # x: a random sample of size 500 from a standard Normal distribution
  y = rnorm(500), # y: a random sample of size 500 from a standard Normal distribution
  log_vec = x+y >1, # A logical vector indicating whether x + y > 1
  num_vec = as.numeric(log_vec),# A numeric vector created by coercing the above logical vector
  factor_vec = as.factor(log_vec)# A factor vector created by coercing the above logical vector
)

```
The size of the data is (`r nrow(df_2) ` ,`r ncol(df_2)`). The mean of x is `r mean(pull(df_2,x)) `. The median of x is `r median(pull(df_2,x))`. The standard deviation of x is `r sd(pull(df_2,x)) `.The proportion of x is `r length(which(pull(df_2,log_vec)))/length(pull(df_2,log_vec))` 

## Promblem 2 
## Make scatterplots
```{r}
ggplot(df_2, aes(x = x, y = y, color = log_vec ))+geom_point() # a scatterplot of y vs x; color points using the logical variable
ggsave("colored_by_logical_variables.pdf") #expoert the image
ggplot(df_2, aes(x = x, y = y, color = num_vec ))+geom_point() 
ggplot(df_2, aes(x = x, y = y, color = factor_vec ))+geom_point() # a second and third scatterplot that color points using the numeric and factor variables, respectively

``` 
Explanation